Verification of concurrent design patterns with data

### 2019-02-07 2018-Arch-pNets/arch-pNets.tex
SMT section, et nouveaux dessins

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{#2}\Simon}}
\newcommand{\noteSBin}[2][inline,color=green!40]{\todo[#1]{{#2}\Simon}}
\newcommand{\todoSB}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Simon:} {#2}}}
\newcommand{\todoSBin}[2][inline,color=green!40]{\todo[#1]{\textbf{To-do Simon: } {#2}}}
\newcommand{\Ludo}{\\\hfill\mdash Ludo}
\newcommand{\noteLH}[2][color=orange!40, size=\tiny]{\todo[#1]{{#2}\Ludo}}
\newcommand{\noteLHin}[2][inline,color=orange!40]{\todo[#1]{{#2}\Ludo}}
\newcommand{\todoLH}[2][color=orange!40, size=\tiny]{\todo[#1]{\textbf{To-do Ludo:} {#2}}}
\newcommand{\todoLHin}[2][inline,color=orange!40]{\todo[#1]{\textbf{To-do Ludo: } {#2}}}
\newcommand{\todoEM}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Eric:} {#2}}}

@@

Parameterised Networks of synchronised automata (pNets) is a formalism for defining behavioural specification of distributed systems.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
In previous work~\cite{HMZ:PDP15}, it was shown pNets can be used to represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods. We
used these results to give the semantics of various constructs and
languages for distributed objects, and to build a platform for design
and verification of distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
By nature, pNets is a  parametrised and hierarchical model.
Their structure is static, but unbounded, and this allows for model-checking
approaches even for reconfigurable applications.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can be used to define composition operators or structuring architectures.
It is possible to reason, in an SMT engine, on the symbolic automaton that represents the behaviour of a pNets with holes and that communicates value~\cite{AVOCS}. This encding of open pNets into Z3 that is under development is the starting point of this article. Our objective is to benefit from the possibility to reason on a pNet in an SMT engine in order to prove properties on BIP architectures.

@@

\begin{indent}
  description courte et intuitive avec reference au papier d'Avocs
\end{indent}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{TimerOADetailed}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{schema:ArchFailure:OA}
\end{figure}

In Figure~\ref{schema:ArchFailure:OA} we show the full open automaton
obtained from our pNet. For reasons of space, we have not shown the
detail of the open transition, but only the assignement of state
variables, and the resulting action; you can find the full details in
Appendix \ref{secn:full-results}. As an example, $OT_3$ is the one
explained in Example~\ref{Example:TimerOT3} page~\pageref{Example:TimerOT3}

This automaton has 12 transitions, including those encoding various
possible firing of some interactions (e.g. OT 2 and 3 for
$fail$). At first sight, it appears the rough behaviour is
reasonable, although a detailed study shows that the property stated
in page~\pageref{property:reset} does not hold, because of the
$fail$ loops on states $11$ and $12$. This is because we did not use
the maximal progress assumption here. If we do, we get the corrected
behaviour in Fig~\ref{schema:ArchFailure:MaxProgress}. This one
verifies all the properties listed in  \app{properties}.

@@

  The new (simplied) OA is in figure \ref{schema:ArchFailure:OA-MaxProgress}.
  Here we give the detail only of the modifed OTs:

@@

\begin{itemize}
\item OT:2 has changed because $b0=true$ is no longer possible
\item OT7 has disappeared because it is now unsat.
\end{itemize}

>>>>

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{#2}\Simon}}
\newcommand{\noteSBin}[2][inline,color=green!40]{\todo[#1]{{#2}\Simon}}
\newcommand{\todoSB}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Simon:} {#2}}}
\newcommand{\todoSBin}[2][inline,color=green!40]{\todo[#1]{\textbf{To-do Simon: } {#2}}}
\newcommand{\Ludo}{\\\hfill\mdash Ludo}
\newcommand{\noteLH}[2][color=orange!40, size=\tiny]{\todo[#1]{{#2}\Ludo}}
\newcommand{\noteLHin}[2][inline,color=orange!40]{\todo[#1]{{#2}\Ludo}}
\newcommand{\todoLH}[2][color=orange!40, size=\tiny]{\todo[#1]{\textbf{To-do Ludo:} {#2}}}
\newcommand{\todoLHin}[2][inline,color=orange!40]{\todo[#1]{\textbf{To-do Ludo: } {#2}}}
\newcommand{\todoEM}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Eric:} {#2}}}
\newcommand{\noteEMin}[2][inline,color=green!40]{\todo[#1]{{#2}\Eric}}

@@

Parameterised Networks of synchronised automata (pNets) is a formalism for defining behavioural specification of distributed systems.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
In previous work~\cite{HMZ:PDP15}, it was shown pNets can be used to represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods. We
used these results to give the semantics of various constructs and
languages for distributed objects, and to build a platform for design
and verification of distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
By nature, pNets is a  parametrised and hierarchical model.
Their structure is static, but unbounded, and this allows for model-checking
approaches even for reconfigurable applications.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can be used to define composition operators or structuring architectures.
It is possible to reason, in an SMT engine, on the symbolic automaton that represents the behaviour of a pNets with holes and that communicates value~\cite{HMZ-FORTE2016QBMZ-AVOCS18}. This encding of open pNets into Z3 that is under development is the starting point of this article. Our objective is to benefit from the possibility to reason on a pNet in an SMT engine in order to prove properties on BIP architectures.

@@

In \cite{HMZ-FORTE2016QBMZ-AVOCS18}, we have described an algorithm
and a tool that 
computes the open automaton encoding the semantics of an open
pNet. This tool computes open transitions from the combinations of all
possible behaviours and synchronisation vectors of the pNet, then makes
use of an SMT engine to check satisfiability of the open transition
predicates, and minimize the size of the resulting automaton.

The encoding into the SMT language involves the definition of the
action and data domains and operators, and the axiomatisation of
domain-dependant properties, to allow the SMT engine to be as precise
as possible (even if it cannot be complete).

The resulting open automaton will later be subject to (symbolic)
model-checking and equivalence checking algorithms, themselves
requiring automatic proof assistance. \todoEM{trop flou pour le
  moment, mais je n'ose pas dire des trucs trop precis... meme si on a
  deja une implem de la FH-bisim, elle n'est pas validee!}

\noteEMin{Est-ce qu'on a la place de mettre un exemple de script SMT ici ?}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{TimerOADetailed}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{schema:ArchFailure:OA}
\end{figure}

In Figure~\ref{schema:ArchFailure:OA} we show the full open automaton
obtained from our pNet. For reasons of space, we have not shown the
detail of the open transition, but only the assignments of state
variables, some usefull parts of the predicates, and the resulting
action; you can find the full details in 
Appendix \ref{secn:full-results}. As an example, $OT_3$ is the one
explained in Example~\ref{Example:TimerOT3} page~\pageref{Example:TimerOT3}

This automaton has 12 transitions, including those encoding various
possible firing of some interactions (e.g. OT 2 and 3 for
$fail$). At first sight, it appears the rough behaviour is
reasonable, although a detailed study shows that the property stated
in page~\pageref{property:reset} does not hold, because of the
$fail$ loops on states $11$ and $12$. This is because we did not use
the maximal progress assumption here. If we do, we get the corrected
behaviour in Fig~\ref{schema:ArchFailure:OA-MaxProgress}
(\app{full-results}), in which $OT_7$ has 
disappeared, and $OT_2$ and $OT_{10}$ have been restricted to
$b0=false$. This one verifies all the properties listed in
\app{properties}.

@@

  The new (simplied) OA is in figure \ref{schema:ArchFailure:OA-MaxProgress}.
  Here we give the detail only of one the modifed OTs:

@@

\begin{itemize}
\item $OT_2$ and $OT_{10}$ have changed because $b0=true$ is no longer possible
\item $OT_7$ has disappeared because it is now unsat.
\end{itemize}
